A tutorial of Bayesian beta regressions with brms in R

### 2024-12-14 manuscript/manuscript.qmd
edit manuscript

Finally, the package marginaleffects has two other convenience functions that return CrIs of comparisons across predictor levels (`avg_comparisons()`) and CrIs of posterior predictions across predictor levels (`avg_predictions`). \[XXX\]

@@

@fig-nas-prop-r plots the posterior distributions of the expected log-odd difference of coarticulatory nasalisation in voiceless vs voiced NC sequences (*x*-axis), for each speaker in the data (*y*-axis), as predicted by the model. The red solid vertical line indicates the constant (overall) expected log-odd difference based on the (constant) `b_` terms. The black dashed vertical line marks log-difference 0 (i.e., no difference in proportion of nasalisation between voiceless and voiced NC). There is a lot of uncertainty within and between speakers: while the distributions of most speakers are located in the positive range, some expected distributions (see bottom of figure) do span both negative and positive values. In other words, while most speakers are expected to have a larger nasalisation proportion in voiceless NC sequences, a few in fact might have the opposite pattern. Even among those speakers that do have a positive difference, there is a lot of uncertainty as to the magnitude of the difference.

@@

>>>>

Finally, the package marginaleffects [XXX] has two other convenience functions that return CrIs of comparisons across predictor levels (`avg_comparisons()`) and CrIs of posterior predictions across predictor levels (`avg_predictions`). \[XXX\]

@@

@fig-nas-prop-r plots the posterior distributions of the expected log-odd difference of coarticulatory nasalisation in voiceless vs voiced NC sequences (*x*-axis), for each speaker in the data (*y*-axis), as predicted by the model. The red solid vertical line indicates the constant (overall) expected log-odd difference based on the (constant) `b_` terms. The black dashed vertical line marks log-difference 0 (i.e., no difference in proportion of nasalisation between voiceless and voiced NC).

@@

There is a lot of uncertainty within and between speakers: while the distributions of most speakers are located in the positive range, some expected distributions (see bottom of figure) do span both negative and positive values. In other words, while most speakers are expected to have a larger nasalisation proportion in voiceless NC sequences, a few in fact might have the opposite pattern. Even among those speakers that do have a positive difference, there is a lot of uncertainty as to the magnitude of the difference.

### 2024-12-14 manuscript/manuscript.tex
edit manuscript

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A tutorial of Bayesian beta regressions with brms in R},
  pdfauthor={Stefano Coretta},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{A tutorial of Bayesian beta regressions with brms in R}
\author[1]{Stefano Coretta%
\corref{cor1}%
\fnref{fn1}}
 \ead{s.coretta@ed.ac.uk}

@@

The \texttt{summary()} function prints the full model summary. For
conciseness, we will use the \texttt{fixef()} function which prints the
regression coefficients. The full summary with an explanation of each
part can be found in XXX. For each coefficient in the model,
\texttt{fixef()} prints the name of the coefficient, the mean estimate,
the estimate error and the lower and upper limits of a Bayesian Credible
interval (CrI). Here, we print an 80\% CrI. There is nothing special
about 95\% CrI within Bayesian inference and instead experts recommend
to check and report a variety of CrIs. Bayesian CrIs indicate that at a
certain probability levels the ``true'' estimate lies within that
interval: so, for example, a 90\% CrI {[}A, B{]} indicates that there is
a 90\% probability that the ``true'' estimate is between A and B.
Different probability levels correspond to different levels of
confidence: the higher the probability the higher the confidence (always
conditional on data and model). Obtaining CrIs at different probability
levels allows researchers to make more fine-grained inferential
statements than the frequentist significance dichotomy affords. For
simplicity of exposition, we will use 80\% CrIs in this case study but
we strongly recommend researchers to always obtain CrIs at different
levels of probability and base their inferences on all and not one in
particular. To reiterate, in Bayesian inference, an 80\% CrI indicates
the range of values within which the true estimate falls at 80\%
probability or confidence. We round all values to the nearest 2 digits
for clarity.

@@

\includegraphics{manuscript_files/figure-pdf/voi-prop-bm-cond-1.pdf}

Finally, the package marginaleffects has two other convenience functions
that return CrIs of comparisons across predictor levels
(\texttt{avg\_comparisons()}) and CrIs of posterior predictions across
predictor levels (\texttt{avg\_predictions}). {[}XXX{]}

@@

\renewcommand\refname{Case study 2: pre-nasalisation in Greek}
  \bibliography{bibliography.bib}

>>>>

\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A tutorial of Bayesian beta regressions with brms in R},
  pdfauthor={Stefano Coretta; Paul Bürkner},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{frontmatter}
\title{A tutorial of Bayesian beta regressions with brms in R}
\author[1]{Stefano Coretta%
\corref{cor1}%
\fnref{fn1}}
 \ead{s.coretta@ed.ac.uk} 
\author[]{Paul Bürkner%
%
}
 \ead{paul.buerkner@gmail.com}

@@

The \texttt{summary()} function prints the full model summary. For
conciseness, we will use the \texttt{fixef()} function which prints the
regression coefficients. {[}EXPLAIN EXPECTED PREDICTIONS XXX{]}. The
full summary with an explanation of each part can be found in XXX. For
each coefficient in the model, \texttt{fixef()} prints the name of the
coefficient, the mean estimate, the estimate error and the lower and
upper limits of a Bayesian Credible interval (CrI). Here, we print an
80\% CrI. There is nothing special about 95\% CrI within Bayesian
inference and instead experts recommend to check and report a variety of
CrIs. Bayesian CrIs indicate that at a certain probability levels the
``true'' estimate lies within that interval: so, for example, a 90\% CrI
{[}A, B{]} indicates that there is a 90\% probability that the ``true''
estimate is between A and B. Different probability levels correspond to
different levels of confidence: the higher the probability the higher
the confidence (always conditional on data and model). Obtaining CrIs at
different probability levels allows researchers to make more
fine-grained inferential statements than the frequentist significance
dichotomy affords. For simplicity of exposition, we will use 80\% CrIs
in this case study but we strongly recommend researchers to always
obtain CrIs at different levels of probability and base their inferences
on all and not one in particular. To reiterate, in Bayesian inference,
an 80\% CrI indicates the range of values within which the true estimate
falls at 80\% probability or confidence. We round all values to the
nearest 2 digits for clarity.

@@

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/voi-prop-bm-cond-1.pdf}}

Finally, the package marginaleffects {[}XXX{]} has two other convenience
functions that return CrIs of comparisons across predictor levels
(\texttt{avg\_comparisons()}) and CrIs of posterior predictions across
predictor levels (\texttt{avg\_predictions}). {[}XXX{]}

@@

\section{Case study 2: coarticulatory vowel
nasalisation}\label{case-study-2-coarticulatory-vowel-nasalisation}

For the second case study we will use data from \citet{carignan2021}.
The study looked at properties of nasality in German VNC sequences.
Here, we will focus on the effect of C voicing (voiceless /t/ vs voiced
/d/) on the proportion of nasalisation within the vowel in the VNC
sequence. Previous work on coarticulatory nasalisation in English has
suggested that vowels followed by an NC sequence where C is voiceless
(NT) should show earlier coarticulatory nasalisation than vowels
followed by an NC sequence where C is voiced \citep[ND, see review
in][]{carignan2021}. This pattern has been suggested to be driven by the
articulatory and acoustic incompatibility of voicelessness and
nasalisation, by which the velum opening gesture of the nasal consonant
is pushed away (i.e.~earlier) when the consonant following the nasal is
voiceless. Everything else being equal, a greater proportion of vowel
nasalisation (from the perspective of time) should be found in vowels
followed by NT than in vowels followed by ND.

We will model the proportion of coarticulatory nasalisation in the
German short vowels /i, e, a, o, u/ when followed by /nt/ or /nd/, using
a Bayesian beta regression model. The proportion was calculated as the
proportion of the nasal interval to the duration of the vowel. The nasal
interval was defined thus: the interval between the time of peak
velocity of velum opening to the offset of the vowel.

First, let's read and plot the data.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nasal }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"data/carignan2021/nasal.csv"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Rows: 705 Columns: 6
-- Column specification --------------------------------------------------------
Delimiter: ","
chr (5): speaker, label, vowel, NC, voicing
dbl (1): nas_prop

i Use `spec()` to retrieve the full column specification for this data.
i Specify the column types or set `show_col_types = FALSE` to quiet this message.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nasal}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 705 x 6
   speaker label               vowel NC    voicing   nas_prop
   <chr>   <chr>               <chr> <chr> <chr>        <dbl>
 1 S03     b__U_nt_@___N_B17/s u     nt    voiceless  0.367  
 2 S03     b__a_nd_@___N_B19/s a     nd    voiced     0.195  
 3 S03     b__a_nt_@___N_B15/s a     nt    voiceless  0.279  
 4 S03     f__I_nt_@___N_B05/s i     nt    voiceless  0.764  
 5 S03     l__I_nd_@___N_B06/s i     nd    voiced     0.00529
 6 S03     p__E_nt_____N_B09/s e     nt    voiceless  0.335  
 7 S03     r__a_nt_@___N_B06/s a     nt    voiceless  0.243  
 8 S03     v__I_nd_@___N_B07/s i     nd    voiced     0.0248 
 9 S03     v__I_nt_6___N_B15/s i     nt    voiceless  0.135  
10 S03     z__E_nd_@___N_B17/s e     nd    voiced     0.538  
# i 695 more rows
\end{verbatim}

\begin{itemize}
\item
  \texttt{speaker} indicates the speaker ID.
\item
  \texttt{label} is the word label as given in the original data.
\item
  \texttt{vowel} is the target vowel in the VNC sequence.
\item
  \texttt{NC} is the NC sequence.
\item
  \texttt{voicing} indicates the voicing of C.
\item
  \texttt{nas\_prop} is the proportion of coarticulatory nasalisation of
  the vowel.
\end{itemize}

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nasal-1.pdf}}

\caption{\label{fig-nasal}Proportion of coarticulatory nasalisation
during the vowel in VNC sequences in German, depending on C voicing.}

Figure~\ref{fig-nasal} shows the proportion of coarticulatory
nasalisation in vowels followed by /nd/ (voiced) vs /nt/ (voiceless)
sequences, for the short vowels /i, e, a, o, u/. We can see a pattern of
higher nasalisation proportion in vowels followed by /nt/, at least in
the vowels /a, i, o/. For /e, u/, the distribution of nasalisation
proportion seems to be similar between the voiced and voiceless
contexts.

Now onto modelling with a beta regression. Note that a full appropriate
model would include further predictors (both constant and varying), but
for simplicity here we include only the following predictors: voicing
(voiced /nd/ vs voiceless /nt/), vowel (/i, e, a, o, u/), including an
interaction between them. As varying terms, we include a varying
intercept by speaker and a by-speaker varying slope for voicing and
vowel in interaction. As with the model from the first case study,
voicing and vowel are coded using indexing, by suppressing the intercept
with \texttt{0\ +}. Here's the code of the model.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
\NormalTok{  nas\_prop }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ voicing}\SpecialCharTok{:}\NormalTok{vowel }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ voicing}\SpecialCharTok{:}\NormalTok{vowel }\SpecialCharTok{|}\NormalTok{ speaker),}
  \AttributeTok{data =}\NormalTok{ nasal,}
  \AttributeTok{family =}\NormalTok{ Beta,}
  \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
  \AttributeTok{seed =} \DecValTok{3749}\NormalTok{,}
  \AttributeTok{file =} \StringTok{"data/cache/nas\_prop\_bm"}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Let's inspect the output of \texttt{fixef()}.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{round}\NormalTok{(}
  \FunctionTok{fixef}\NormalTok{(nas\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
  \AttributeTok{digits =} \DecValTok{2}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
                        Estimate Est.Error   Q10   Q90
voicingvoiced:vowela       -0.61      0.12 -0.77 -0.46
voicingvoiceless:vowela    -0.08      0.09 -0.19  0.04
voicingvoiced:vowele       -0.19      0.16 -0.38  0.01
voicingvoiceless:vowele    -0.25      0.10 -0.38 -0.12
voicingvoiced:voweli       -0.74      0.19 -0.98 -0.51
voicingvoiceless:voweli    -0.12      0.18 -0.35  0.10
voicingvoiced:vowelo       -0.34      0.16 -0.54 -0.14
voicingvoiceless:vowelo     0.25      0.15  0.05  0.44
voicingvoiced:vowelu       -0.71      0.18 -0.94 -0.49
voicingvoiceless:vowelu    -0.58      0.16 -0.78 -0.38
\end{verbatim}

Negative log-odds indicate a proportion that is smaller than 50\%, while
positive log-odds a proportion that is greater than 50\%. Generally, the
expected log-odd predictions are negative, indicating an overall
tendency for the nasalisation to take less than 50\% of the duration of
the vowel. Moreover, the predictions are higher for voiceless NC
sequences than for voiced NC sequences, indicating a greater proportion
of nasalisation in the former. However there is vowel-specific
variation, and there doesn't seem to be much of a difference in
nasalisation proportion in /e/ and /u/. Plotting the expected
predictions with \texttt{conditional\_effects()} should make this
clearer.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{conditional\_effects}\NormalTok{(nas\_prop\_bm, }\StringTok{"vowel:voicing"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/nas-prop-1.pdf}}

Now that we fitted the model we can use the draws to answer two
questions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Is the nasalisation proportion, on average across vowels, greater in
  voiceless NC sequences?
\item
  Is there individual speaker variation?
\end{enumerate}

To answer question 1, we can calculate the average difference in
nasalisation proportion by first calculating the average nasalisation
across all vowels for voiced and voiceless sequences and then take the
difference of those, similarly to what we have done in the Case Study 1
above.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# extract only coefficient variables starting with "b\_"}
\NormalTok{nas\_prop\_bm\_draws }\OtherTok{\textless{}{-}} \FunctionTok{as\_draws\_df}\NormalTok{(nas\_prop\_bm, }\AttributeTok{variable =} \StringTok{"\^{}b\_"}\NormalTok{, }\AttributeTok{regex =} \ConstantTok{TRUE}\NormalTok{)}

\NormalTok{nas\_prop\_bm\_draws\_long }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot vowel:c2 columns}
  \FunctionTok{pivot\_longer}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{b\_voicingvoiced:vowela}\StringTok{\textasciigrave{}}\SpecialCharTok{:}\StringTok{\textasciigrave{}}\AttributeTok{b\_voicingvoiceless:vowelu}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"coeff"}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
  \CommentTok{\# separate "coeff" labels into type ("b"), vowel and c2}
  \FunctionTok{separate}\NormalTok{(coeff, }\AttributeTok{into =} \FunctionTok{c}\NormalTok{(}\StringTok{"type"}\NormalTok{, }\StringTok{"voicing"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_draws\_long}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 40,000 x 7
   .chain .iteration .draw type  voicing          vowel    value
    <int>      <int> <int> <chr> <chr>            <chr>    <dbl>
 1      1          1     1 b     voicingvoiced    vowela -0.463 
 2      1          1     1 b     voicingvoiceless vowela -0.0944
 3      1          1     1 b     voicingvoiced    vowele -0.155 
 4      1          1     1 b     voicingvoiceless vowele -0.208 
 5      1          1     1 b     voicingvoiced    voweli -0.481 
 6      1          1     1 b     voicingvoiceless voweli -0.0529
 7      1          1     1 b     voicingvoiced    vowelo -0.310 
 8      1          1     1 b     voicingvoiceless vowelo  0.321 
 9      1          1     1 b     voicingvoiced    vowelu -0.620 
10      1          1     1 b     voicingvoiceless vowelu -0.368 
# i 39,990 more rows
\end{verbatim}

Now let's calculate the mean nasalisation proportion within each draw by
voicing, and plot the resulting posterior distributions.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_draws\_long\_voicing }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# grouping by .draw and voicing ensures that averaging applies only within draw and voicing}
  \FunctionTok{group\_by}\NormalTok{(.draw, voicing) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# calculate the mean value within draw/voicing}
  \FunctionTok{summarise}\NormalTok{(}
    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value), }\AttributeTok{.groups =} \StringTok{"drop"}
\NormalTok{  )}

\NormalTok{nas\_prop\_bm\_draws\_long\_voicing }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value\_mean), voicing)) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/nas-prop-bm-draws-long-voicing-1.pdf}}

The plot suggests an overall greater nasalisation proportion in
voiceless NC sequences. Let's quantify how greater as we did in Case
Study 1.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_bm\_diff }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws\_long\_voicing }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# pivot data to create one column per voicing with the mean drawn values,}
  \CommentTok{\# with one draw per raw}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ voicing, }\AttributeTok{values\_from =}\NormalTok{ value\_mean) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}
    \CommentTok{\# calculate the difference of voiceless and voiced}
    \AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced}
\NormalTok{  )}

\NormalTok{nas\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{reframe}\NormalTok{(}
    \CommentTok{\# 90\% CrI}
    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(}\FunctionTok{exp}\NormalTok{(voicing\_diff), }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
    \CommentTok{\# 80\% CrI}
    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(}\FunctionTok{exp}\NormalTok{(voicing\_diff), }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
    \CommentTok{\# 60\% CrI}
    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(}\FunctionTok{exp}\NormalTok{(voicing\_diff), }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
\NormalTok{  ) }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{round}\NormalTok{(.x, }\DecValTok{2}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 2 x 3
    q90   q80   q60
  <dbl> <dbl> <dbl>
1  1.23  1.27  1.33
2  1.69  1.63  1.56
\end{verbatim}

The CrIs of the ratio difference in nasalisation proportion in voiceless
vs voiced NC sequences suggest a robust increase of nasalisation in the
voiceless NC sequences, with a 90\% probability that the increase is
between 23\% and 69\% of the proportion in voiced NC sequences.

Moving onto question 2: is there individual speaker variation?
{[}TAMMINGA XXX{]} For this, we will use the \texttt{spread\_draws()}
function from tidybayes {[}XXX{]} to extract the draws of the varying
terms (in brms these are the coefficients that start with \texttt{r\_}).
There is quite a few steps of processing to get from the raw draws to
the estimand we need: while we have commented the following code, we
encourage readers to test each line sequentialy and inspect the
intermediate output to fully understand the process.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(tidybayes)}
\end{Highlighting}
\end{Shaded}

Attaching package: 'tidybayes'
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_r }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# extract draws from model, only \textasciigrave{}r\_speaker\textasciigrave{} varying terms}
  \FunctionTok{spread\_draws}\NormalTok{(r\_speaker[speaker,voicingvowel]) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# separate the column voicingvowel to two columns}
  \FunctionTok{separate}\NormalTok{(voicingvowel, }\FunctionTok{c}\NormalTok{(}\StringTok{"voicing"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# join the draws with the \textasciigrave{}b\_\textasciigrave{} terms}
  \FunctionTok{left\_join}\NormalTok{(}\AttributeTok{y =}\NormalTok{ nas\_prop\_bm\_draws\_long) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# get the expected log{-}odd value of each speaker, in each draw}
  \CommentTok{\# this is the sum of the \textasciigrave{}value\textasciigrave{} from the b\_ terms and the value from the}
  \CommentTok{\# r\_speaker term.}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{r\_speaker\_value =}\NormalTok{ value }\SpecialCharTok{+}\NormalTok{ r\_speaker) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# group the data for summarise}
  \FunctionTok{group\_by}\NormalTok{(.draw, speaker, voicing) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# get mean expected log{-}odds by draw, speaker and voicing (averaging across vowel)}
  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{r\_speaker\_value\_mean =} \FunctionTok{mean}\NormalTok{(r\_speaker\_value)) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# make the data wider: two columns, one for voiced and one for voiceless}
  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ voicing, }\AttributeTok{values\_from =}\NormalTok{ r\_speaker\_value\_mean) }\SpecialCharTok{|\textgreater{}} 
  \CommentTok{\# finally, calculate the difference in expected log{-}odds of voiceless and voiced}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Joining with `by = join_by(voicing, vowel, .chain, .iteration, .draw)`
\end{verbatim}

\begin{verbatim}
`summarise()` has grouped output by '.draw', 'speaker'. You can override using
the `.groups` argument.
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_r}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
# A tibble: 140,000 x 5
# Groups:   .draw, speaker [140,000]
   .draw speaker voicingvoiced voicingvoiceless voicing_diff
   <int> <chr>           <dbl>            <dbl>        <dbl>
 1     1 S03          -1.42              0.0676      1.49   
 2     1 S04           0.209             0.185      -0.0238 
 3     1 S05          -0.284            -0.0546      0.229  
 4     1 S06           0.00669          -0.324      -0.331  
 5     1 S07          -0.446            -0.147       0.299  
 6     1 S08          -0.610            -0.127       0.483  
 7     1 S09          -0.473             0.0497      0.523  
 8     1 S10          -0.556            -0.311       0.245  
 9     1 S11           0.201            -0.426      -0.628  
10     1 S12          -0.0812           -0.0856     -0.00439
# i 139,990 more rows
\end{verbatim}

Figure~\ref{fig-nas-prop-r} plots the posterior distributions of the
expected log-odd difference of coarticulatory nasalisation in voiceless
vs voiced NC sequences (\emph{x}-axis), for each speaker in the data
(\emph{y}-axis), as predicted by the model. The red solid vertical line
indicates the constant (overall) expected log-odd difference based on
the (constant) \texttt{b\_} terms. The black dashed vertical line marks
log-difference 0 (i.e., no difference in proportion of nasalisation
between voiceless and voiced NC).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nas\_prop\_r }\SpecialCharTok{|\textgreater{}} 
  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(voicing\_diff, }\FunctionTok{reorder}\NormalTok{(speaker, voicing\_diff))) }\SpecialCharTok{+}
  \FunctionTok{stat\_halfeye}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \FunctionTok{mean}\NormalTok{(nas\_prop\_bm\_diff}\SpecialCharTok{$}\NormalTok{voicing\_diff), }\AttributeTok{colour =} \StringTok{"red"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{lims}\NormalTok{(}\AttributeTok{x =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\FloatTok{1.5}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nas-prop-r-1.pdf}}

\caption{\label{fig-nas-prop-r}}

There is a lot of uncertainty within and between speakers: while the
distributions of most speakers are located in the positive range, some
expected distributions (see bottom of figure) do span both negative and
positive values. In other words, while most speakers are expected to
have a larger nasalisation proportion in voiceless NC sequences, a few
in fact might have the opposite pattern. Even among those speakers that
do have a positive difference, there is a lot of uncertainty as to the
magnitude of the difference.

  \bibliography{linguistics.bib}
